---
title: "Assigment1"
author: "Li"
date: "December 20, 2015"
output: html_document
---

Loading the data

```{r}
activity <- read.csv(file = "activity.csv")
```
Q1: What is mean total number of steps taken per day?
=====================================================
Format the dataset as per day
```{r}
library(reshape2)
activity_perday <- melt(activity, id = "date")
activity_perday <- dcast(activity_perday, date~variable, sum)
```

Make histogram of the total number of steps taken each day

```{r}
hist(activity_perday$steps, main="Histogram of total number of steps taken each day", xlab = "steps", breaks = 5)
```

Calculate the mean and median of the total number of steps taken per day

```{r}
mean(activity_perday$steps, na.rm = TRUE)
median(activity_perday$steps, na.rm = TRUE)
```

Q2: What is the average daily activity pattern?
===============================================
Make the time series plot
First, format the dataset by group steps by interval. Then make a plot.

```{r}
library(dplyr)
byinterval <- group_by(activity, interval)
q2plotdata <- summarise(byinterval, stepmean = mean(steps, na.rm = T))
with(q2plotdata, {plot(interval, stepmean, type = 'l', xlab = "time interval", ylab="mean steps", main="The time series plot of average number of steps taken")})
```

Find out which time interval contains the maximum number of steps.
```{r}
q2plotdata[which.max(q2plotdata$stepmean),]
```

Q3 Imputing missing values
==========================
Calculate and report the total number of missing values in the dataset

```{r}
sapply(activity, function(x) sum(is.na(x)))
```

Fill the missing value with the mean of that 5-minite interval.
```{r}
activity$steps2 <- activity$steps
naplace <- which(is.na(activity$steps))
for(i in naplace){
  j <- i%%288
  if(j != 0) {
    activity$steps2[i] <- q2plotdata$stepmean[j]
  } else{
    activity$steps2[i] <- q2plotdata$stepmean[288]
  }  
}
```

Create a new dataset that is equal to the original dataset with non missing value.

```{r}
activity2 <- activity[, 2:4]
sapply(activity2, function(x) sum(is.na(x)))
summary(activity2)
```

Plot the total number of steps taken each day, and caculate mean and median from the new dataset.

```{r}
activity2perday <- melt(activity2, id = "date")
activity2perday <- dcast(activity2perday, date~variable, sum)
hist(activity2perday$steps, main="Histogram of total number of steps taken each day with missing value filled", xlab = "steps", breaks = 5)
mean(activity2perday$steps, na.rm = TRUE)
median(activity2perday$steps, na.rm = TRUE)
```

Q4 Are there differences in activity patterns between weekdays and weekends?
============================================================================
Translate the date as a factor

```{r}
activity2$day <- weekdays(as.Date(activity2$date))
weekend <- c("Saturday", "Sunday")
activity2$week <- factor((activity2$day %in% weekend), levels = c(TRUE, FALSE), labels=c('weekend','weekday'))
byinterval2 <- group_by(activity2, interval, week)
q4plotdata <- summarise(byinterval2, stepmean = mean(steps2))
````

Make plot to show difference between weekday and weekend
```{r}
library(ggplot2)
qplot(interval, stepmean, data= q4plotdata, geom="line", xlab="time interval", ylab = "mean steps", facets = week~.)
```


